En garde! Unguarded iteration for reversible computation in the Delay monad\thanks{Supported by organization x.}

### 2019-04-29 mpc19/paper.tex
soundness
Added diagram for Hasegawa trace.

\section{TODO} % (fold)
\label{sec:todo}
Robin:
\begin{itemize}
  \item Make pictures! 
  \begin{itemize}
    \item Hasegawa trace from iterator.
  \end{itemize}
  \item Conclusion
  \begin{itemize}
    \item Related work: Kleisli category of the Delay monad is a join
    restriction category, inverse category of a join restriction category is a
    join inverse category; already known to be dagger-traced under certain
    conditions.
    \item Future work: Semantics of time invertible programming. Extension of
    the work on the Kleisli category of the Delay monad beyond $\Set$ to other
    (more general) categories, e.g. topoi.
  \end{itemize}
\end{itemize}
Niccolò: 
\begin{itemize}
  \item Fix the syntax to be closer to ordinary presentations (but do describe 
  how it is realized in Agda).
  \item Finish the section about trace and iteration.
  \item Remind the reader about double lines of coinductive definitions.
  \item Swap types in Elgot iterator.
\end{itemize}
Both:
\begin{itemize}
  \item \textbf{Important:} Come up with a punnier title! (Something about 
  delays and daggers).
\end{itemize}
% section todo (end)

@@

It is well-known that a trace operator is obtainable from an iteration
operator, as shown by Hasegawa  \cite{Hasegawa97}. His construction,
instantiated to our setting, looks as follows:
\begin{align*}
& \traceHD : (A + C \to \Delay (B + C)) \to A \to \Delay B \\
& \traceHD \,f = \iterD (\Delay (\id + \inr) \comp f) \comp \inl
\end{align*}

It is not difficoult to prove that the two possible ways of defining a
trace operator from Elgot iteration are equivalent, in the sense that
$\traceD\,f \approx \traceHD\,f$ for all $f : A + C \to \Delay (B +
C)$.

@@

The notion of orbit can be used to state when 
the iteration of a function $f : A \to \Delay (B + A)$ on a input
$a : A$ terminates with value $b : B$.
\[
\iterD\,f\,a \dn b \leftrightarrow \Orb {\copairD{\inlD}{f}} {(\inr\,a)} {(\inl\,b)}
\]
We refer the intersted reader to our Agda formalization for a complete
proof of this logical equivalence.
Similarly, the orbit can be used to state when 
the trace of a function $f : A + C \to \Delay (B + C)$ on a input
$a : A$ terminates with value $b : B$.
\begin{equation}\label{eq:traceorbit}
\traceD\,f\,a \dn b \leftrightarrow \Orb f{(\inl\,a)} {(\inl\,b)}
\end{equation}

>>>>

\section{TODO} % (fold)
\label{sec:todo}
Robin:
\begin{itemize}
  \item Conclusion
  \begin{itemize}
    \item Related work: Kleisli category of the Delay monad is a join
    restriction category, inverse category of a join restriction category is a
    join inverse category; already known to be dagger-traced under certain
    conditions.
    \item Future work: Semantics of time invertible programming. Extension of
    the work on the Kleisli category of the Delay monad beyond $\Set$ to other
    (more general) categories, e.g. topoi.
  \end{itemize}
\end{itemize}
Niccolò: 
\begin{itemize}
  \item Fix the syntax to be closer to ordinary presentations (but do describe 
  how it is realized in Agda).
  \item Finish the section about trace and iteration.
  \item Remind the reader about double lines of coinductive definitions.
  \item Swap types in Elgot iterator.
\end{itemize}
Both:
\begin{itemize}
  \item \textbf{Important:} Come up with a punnier title! (Something about 
  delays and daggers).
\end{itemize}
% section todo (end)

@@

It is well-known that a trace operator is obtainable from an iteration
operator, as shown by Hasegawa~\cite{Hasegawa97}. His construction,
instantiated to our setting, looks as follows:
\begin{align*}
& \traceHD : (A + C \to \Delay (B + C)) \to A \to \Delay B \\
& \traceHD \,f = \iterD (\Delay (\id + \inr) \comp f) \comp \inl
\end{align*}
or graphically
\ctikzfig{trace_hasegawa}

It is not difficult to prove that the two possible ways of defining a
trace operator from Elgot iteration are equivalent, in the sense that
$\traceD\,f \approx \traceHD\,f$ for all $f : A + C \to \Delay (B +
C)$.

@@

The notion of orbit can be used to state when 
the iteration of a function $f : A \to \Delay (B + A)$ on a input
$a : A$ terminates with value $b : B$.
\[
\iterD\,f\,a \dn b \leftrightarrow \Orb {\copairD{\inlD}{f}} {(\inr\,a)} {(\inl\,b)}
\]
We refer the interested reader to our Agda formalization for a complete
proof of this logical equivalence.
Similarly, the orbit can be used to state when 
the trace of a function $f : A + C \to \Delay (B + C)$ on a input
$a : A$ terminates with value $b : B$.
\begin{equation}\label{eq:traceorbit}
\traceD\,f\,a \dn b \leftrightarrow \Orb f{(\inl\,a)} {(\inl\,b)}
\end{equation}
